\documentclass[11pt]{article}

%old versions on or before 2003/03/27
\usepackage{apacite}
%new versions of apacite should use this variant
%\usepackage[BCAY]{apacite}

\begin{document}

this topic, including the authors of the second article. 
The other one, written by W. Vialle et al. \cite{vialle2001}
Kulik and Kulik performed a meta-analysis about the effects of acceleration of
gifted children \cite{kulik1984}. The authors state that many older studies

\cite{hooper1986,ortony1988,russell1979} 
\cite<in>[chap.~3]{allen1995a,allen1995b}
\citeA{allen1995a}
\citeyear{allen1995b}
\citeauthor{allen1995b}
\citeyearNP{allen1995a}

\begin{thebibliography}{}


\bibitem[\protect\BCAY{Kulik \BBA{} Kulik}{Kulik \BBA{}
  Kulik}{1984}]{kulik1984}
Kulik, J.~A.\BCBT{} \BBA{} Kulik, C.-L.~C.
\newblock{}\BBOP{}1984\BBCP{}.
\newblock{}\BBOQ{}{Effects of Accelerated Instruction on Students}.\BBCQ{}
\newblock{}\Bem{Review of Educational Research}, \Bem{54}(3).

\bibitem[\protect\BCAY{Vialle, Ashton, Carlon\BCBL{} \BBA{} Rankin}{Vialle
  \BOthers{}}{2001}]{vialle2001}
Vialle, W., Ashton, T., Carlon, G.\BCBL{} \BBA{} Rankin, F.
\newblock{}\BBOP{}2001\BBCP{}.
\newblock{}\BBOQ{}{Acceleration: A Coat of Many Colours}.\BBCQ{}
\newblock{}\Bem{Roeper Review}, \Bem{24}(1).

\bibitem[\protect\BCAY{Allen \BBA{} Carifio}{Allen \BBA{}
  Carifio}{1995\protect\BCnt{1}}]{allen1995b}
Allen, B.~D.\BCBT{} \BBA{} Carifio, J.
\newblock{}\BBOP{}1995\protect\BCnt{1}, May\BBCP{}.
\newblock{}\Bem{The relationship between emotional states and solving complex
  mathematical problems.}
\newblock{}Paper presented at the meeting of Eastern Educational Research
  Association, Hilton Head, SC.

\bibitem[\protect\BCAY{Allen \BBA{} Carifio}{Allen \BBA{}
  Carifio}{1995\protect\BCnt{2}}]{allen1995a}
Allen, B.~D.\BCBT{} \BBA{} Carifio, J.
\newblock{}\BBOP{}1995\protect\BCnt{2}, March\BBCP{}.
\newblock{}\Bem{A {M}ethodology for the {A}nalysis of {E}motion {E}xperiences
  during {M}athematical {P}roblem {S}olving.}
\newblock{}Paper presented at the Annual Conference of the New England
  Educational Research Organization, Portsmouth, NH.

\bibitem[\protect\BCAY{Hooper}{Hooper}{1986}]{hooper1986}
Hooper, J.
\newblock{}\BBOP{}1986\BBCP{}.
\newblock{}\Bem{The 3-{P}ound {U}niverse.}
\newblock{}Los Angeles: Teacher, Inc.

\bibitem[\protect\BCAY{Ortony, Clore\BCBL{} \BBA{} Collins}{Ortony
  \BOthers{}}{1988}]{ortony1988}
Ortony, A., Clore, G.\BCBL{} \BBA{} Collins, A.
\newblock{}\BBOP{}1988\BBCP{}.
\newblock{}\Bem{The {C}ognitive {S}tructure of {E}motions.}
\newblock{}New York: Cambridge University Press.

\bibitem[\protect\BCAY{Russell}{Russell}{1979}]{russell1979}
Russell, J.~A.
\newblock{}\BBOP{}1979\BBCP{}.
\newblock{}\BBOQ{}Affect {S}pace is {B}ipolar.\BBCQ{}
\newblock{}\Bem{Journal of Personality and Social Psychology}, \Bem{37}(3),
  345--356.

\end{thebibliography}
\end{document}
